---
title: "Benthic Reformatting"
author: "Jeanette Clark"
date: "7/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This doc checks the excel file sent by Betsy Wells for potential QA issues.

```{r, message = FALSE}
library(readxl)
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(janitor)
library(sf)
library(ggmap)
```

Reformat the main sheet from wide to long.

```{r, warning = FALSE, message = FALSE}
d <- read_excel(
        "../../data/benthic/1975-Oct2020 DWR EMP Benthic CPUE 20210511.xlsx",
        skip = 7,
        col_names = T,
        .name_repair = 'minimal')

# find columns to not pivot over
fixed_cols <-
    grep("[0-9]", colnames(d), value = TRUE, invert = TRUE)

# read in just the first 6 rows with taxa info
h <- read_excel(
        "../../data/benthic/1975-Oct2020 DWR EMP Benthic CPUE 20210511.xlsx",
        skip = 1,
        n_max = 7,
        col_names = F
    )

# paste all taxa together
col_n <- map_chr(h, paste, collapse = "|")

# replace appropriate column names in main data frame with pasted taxa info
col_n[grep("^NA|NA|NA", col_n)] <- NA
col_d <- colnames(d)
col_d[grep("[0-9]", col_d)] <- col_n[!is.na(col_n)]
colnames(d) <- col_d


# move from wide to long and split taxa info back into columns
cpue <- d %>% pivot_longer(
        cols = -all_of(fixed_cols),
        names_to = c(
            "phylum",
            "class",
            "order",
            "family",
            "genus",
            "species",
            "id_no"),
        values_to = "count",
        names_sep = "\\|") %>%
    mutate_all(as.character) %>% 
    clean_names() %>% 
    unite(species, genus, species, sep = " ") %>% 
    mutate(count = as.numeric(count),
           year = as.numeric(year))
```

```{r}
stations <- read_excel("../../data/benthic/1975-Oct2020 DWR EMP Benthic CPUE 20210511.xlsx", sheet = 2) %>% 
    clean_names() %>% 
    rename(station_code = site_code)
```

Check to make sure all counts have station information

```{r}
nrow(anti_join(cpue, stations, by = "station_code"))
```

Get total site visits per year

```{r}
visits <- read_excel("../../data/benthic/1975-Oct2020 DWR EMP Benthic CPUE 20210511.xlsx", sheet = 4) %>% 
    rename(station_code = `Row Labels`) %>% 
    pivot_longer(-station_code, names_to = "year", values_to = "visits") %>% 
    filter(station_code != "Grand Total" & year != "Grand Total") %>% 
    mutate(year = as.numeric(year)) %>% 
    arrange(station_code, year)
```

Get total grabs per year

```{r}
grabs <- read_excel("../../data/benthic/1975-Oct2020 DWR EMP Benthic CPUE 20210511.xlsx", sheet = 5) %>% 
    rename(station_code = `Row Labels`) %>% 
    pivot_longer(-station_code, names_to = "year", values_to = "grabs") %>% 
    filter(station_code != "Grand Total" & year != "Grand Total") %>% 
    mutate(year = as.numeric(year)) %>% 
    arrange(station_code, year)
```

Check to make sure visits and grabs have valid station information.

```{r}
g_j <- anti_join(grabs, stations, by = "station_code")
v_j <- anti_join(visits, stations, by = "station_code")

nrow(g_j)
nrow(v_j)
```

# Check the invasive species data

```{r}
inv <- cpue %>% 
    filter(species %in% c("Potamocorbula amurensis", "Corbicula fluminea"))
```

```{r}
ggplot(inv, aes(x = year, y = count, color = species)) +
    geom_point() +
    theme_classic()
```

# Map mean invasives at each study location

```{r}
inv_mean <- inv %>% 
    group_by(station_code, species) %>% 
    summarise(count = mean(count), .groups = "drop") %>% 
    left_join(stations, by = "station_code") %>% 
    st_as_sf(coords = c('longitude', 'latitude'),
             crs = 4326)
```

Get ggmap basemap

```{r, message = FALSE}
# Define a function to fix the bbox to be in EPSG:3857
# See https://github.com/dkahle/ggmap/issues/160#issuecomment-397055208
ggmap_bbox_to_3857 <- function(map) {
  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
  # and set the names to what sf::st_bbox expects:
  map_bbox <- setNames(unlist(attr(map, "bb")), 
                       c("ymin", "xmin", "ymax", "xmax"))
  
  # Coonvert the bbox to an sf polygon, transform it to 3857, 
  # and convert back to a bbox (convoluted, but it works)
  bbox_3857 <- st_bbox(st_transform(st_as_sfc(st_bbox(map_bbox, crs = 4326)), 3857))
  
  # Overwrite the bbox of the ggmap object with the transformed coordinates 
  attr(map, "bb")$ll.lat <- bbox_3857["ymin"]
  attr(map, "bb")$ll.lon <- bbox_3857["xmin"]
  attr(map, "bb")$ur.lat <- bbox_3857["ymax"]
  attr(map, "bb")$ur.lon <- bbox_3857["xmax"]
  map
}

bbox <- c(-122.25, 37.78, -121.31, 38.35)   # This is roughly southern Alaska
basemap <- get_stamenmap(bbox, zoom = 10)
basemap_3857 <- ggmap_bbox_to_3857(basemap)
```


```{r, message = FALSE}
inv_mean_3857 <- st_transform(inv_mean, crs = 3857)

ggmap(basemap_3857) +
    geom_sf(data = inv_mean_3857,
            aes(color = species, size = count),
            inherit.aes = FALSE,
            shape = 21,
            stroke = 1) +
    scale_size(range = c(0.5, 12)) +
    scale_color_manual(values = c("#fc8d62", "#377eb8")) +
    labs(x = "", y = "")
```

